---
title: "Inkscape extensions, actions and verbs - advanced"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: false
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{The inkscaper package: R wrapper for Inkscape - advanced}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---




## Inkscape extensions



To build a dataframe with the names and descriptions of the Inkscape extensions on Windows you can use this script:

```{r, eval=FALSE}
inx_extensions_list_win <- function(){
  path = system('inkscape --system-data-directory', intern = TRUE)
  inkscape_extensions_path = paste(path, "\\extensions", sep = "")
  inkscape_python_home  = paste(gsub("\\share\\inkscape", "", path, fixed = T), "\\bin", sep = "")
  inkscape_extension_names = list.files(path = inkscape_extensions_path, pattern = ".py")
  df <- data.frame(Extension = inkscape_extension_names, Description = NA)
  for (i in c(1:nrow(df))) {
    inkscape_extension_name = df$Extension[i]
    print(inkscape_extension_name)
    output = tempfile("inx_", fileext = ".txt")
    con  = tempfile(pattern = "inx_", fileext = ".bat")
    fmt = '@ECHO OFF
cd %s
python.exe "%s" --help  > %s'
    text = sprintf(fmt,
                   inkscape_python_home,
                   paste(inkscape_extensions_path, inkscape_extension_name, sep = "\\"),
                   output)
    writeLines(text, con)
    system(con)
    output
    inkscape_extension_description <- readLines(output)
    df$Description[i] <- paste(inkscape_extension_description, collapse = "\n")
    }

  df
}

inx_extensions_list_win() %>% readr::write_csv("../inst/extdata/extensions.csv")
```

To build a dataframe with the names and descriptions of the Inkscape extensions on Linux you can use this script:

```{r, eval=FALSE}
inx_extensions_list_linux <- function(){
  path = system('inkscape --system-data-directory', intern = TRUE)
  inkscape_extensions_path = paste(path, "/extensions", sep = "")
  inkscape_extension_names = list.files(path = inkscape_extensions_path, pattern = ".py")
  df <- data.frame(Extension = inkscape_extension_names, Description = NA)
  for (i in c(1:nrow(df))) {
    inkscape_extension_name = df$Extension[i]
    print(inkscape_extension_name)
    output = tempfile("inx_", fileext = ".txt")
    command <- sprintf('python %s --help  > %s', paste(inkscape_extensions_path, inkscape_extension_name, sep = "/"), output)
    system(command, intern = TRUE)
    inkscape_extension_description <- readLines(output)
    df$Description[i] <- paste(inkscape_extension_description, collapse = "\n")
  }
  df
}

```

Here is the dataframe with the names and descriptions of the Inkscape (`r system('inkscape --version', intern = TRUE)`) extensions:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
readr::read_csv("../inst/extdata/extensions.csv") %>% kbl() %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")
```


## Inkscape actions

[Actions](https://wiki.inkscape.org/wiki/Using_the_Command_Line) give you the ability to batch SVG files in Inkscape without opening a Graphical User Interface (GUI).

To build a dataframe with the names and descriptions of the Inkscape actions on Windows you can use this script:

```{r, eval=FALSE}
inx_actions_list_win <- function(){
  output = tempfile("inx_", fileext = ".txt")
  con  = tempfile(pattern = "inx_", fileext = ".bat")
  fmt = '@ECHO OFF
inkscape --action-list > %s'
  text = sprintf(fmt, output)
  writeLines(text, con)
  system(con)
  actions <- readr::read_delim(output, delim = ":", col_names = FALSE)
  names(actions) = c("Action", "Description")
  actions
}
inx_actions_list_win() %>% readr::write_csv("../inst/extdata/actions.csv")
```

To build a dataframe with the names and descriptions of the Inkscape actions on Linux you can use this script:

```{r, eval=FALSE}
inx_actions_list_linux <- function(){
  output = tempfile("inx_", fileext = ".txt")
  fmt = 'inkscape --action-list > %s'
  text = sprintf(fmt, output)
  system(text)
  actions <- readr::read_delim(output, delim = ":", col_names = FALSE)
  names(actions) = c("Action", "Description")
  actions
}
```

Here is the dataframe with the names and descriptions of the Inkscape (`r system('inkscape --version', intern = TRUE)`) actions:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
readr::read_csv("../inst/extdata/actions.csv") %>% kbl() %>%
  kable_paper() %>%
  scroll_box(width = "100%", height = "300px")
```


Here is the dataframe with the name, description and GUI requirement indicator of the Inkscape (`r system('inkscape --version', intern = TRUE)`) verbs:

