---
title: "The magick package: Advanced Image-Processing in R"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: false
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{The magick package: Advanced Image-Processing in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
dev.off <- function(){
  invisible(grDevices::dev.off())
}
```

The [magick](https://cran.r-project.org/package=magick) package provide a modern and simple toolkit for image processing in R. It wraps the [ImageMagick STL](https://www.imagemagick.org/Magick++/STL.html) which is the most comprehensive open-source image processing library available today.

The ImageMagick library has an overwhelming amount of functionality. Magick exposes a decent subset of it, but it is impossible to document everything in detail. This article introduces some basic concepts and examples to get started.

## Installing `magick`

installing on windows - add path

On Windows or macOS the package is most easily installed via CRAN.

```r
#install.packages("magick")
```

The binary CRAN packages work out of the box and have most important features enabled.
Use `magick_config` to see which features and formats are supported by your version of ImageMagick.


```{r}

```


### Examples

